### 2023-05-15 Rmarkdown/paper-si.Rmd
first commit

---
title: "*Supplementary information for:* Sentinel pipe racks quantify orthophosphate's effect on lead release into drinking water"
author: "Benjamin F. Trueman^§,^\\*, Wendy H. Krkošek^†^, and Graham A. Gagnon^§^"
date: "^§^Centre for Water Resources Studies, Department of Civil & Resource Engineering, Dalhousie University, 1360 Barrington St., Halifax, Nova Scotia, Canada B3H 4R2  \n&nbsp;  \n^†^Halifax Water, 450 Cowie Hill Rd., Halifax, Nova Scotia, Canada,  B3P 2V3  \n&nbsp;   \n&nbsp;*Corresponding author  \n&nbsp;E-mail: benjamin.trueman@dal.ca    \n&nbsp;Tel:902.494.6070  \n&nbsp;Fax: 902.494.3105"
output:
  word_document:
    reference_docx: StylesTemplate.docx
bibliography: references.bib
csl: environmental-science-and-technology.csl
---

```{r setup, include=FALSE}
here::i_am("Rmarkdown/paper.Rmd")
knitr::opts_chunk$set(echo = FALSE, dev = "png", dpi = 600)
options(dplyr.summarise.inform = FALSE)
source(here::here("R/03-inform-prior-ml.R"))
source(here::here("R/04-model-ml.R"))
library("posterior")
wq <- read_csv(here::here("data-clean/summary-table.csv"))
wq_si <- read_csv(here::here("data-clean/summary-table-si.csv"))
```

```{r functions}
# this is the opposite of fill()

unfill_vec <- function(x, placeholder = "") {
  same <- x == dplyr::lag(x)
  ifelse(!is.na(same) & same, placeholder, as.character(x))
}

```

This document has 5 pages, 5 figures, and 1 table.

```{r summary-wq-si}
options(knitr.kable.NA = "")
wq_si %>% 
  transmute(
    Parameter = unfill_vec(param),
    Unit = unfill_vec(str_replace(unit, "ug", "µg")),
    Zone = zone,
    Median = round(med, 1), 
    `Lower quartile` = round(lq, 1),
    `Upper quartile` = round(uq, 1)
  ) %>% 
  knitr::kable(caption = "**Table S1.** Summary of water quality in pipe rack effluent, by zone.")

```

```{r figure-s1, fig.cap="**Figure S1.** An example of the pipe racks installed in Zones 1 and 2."}
knitr::include_graphics(here::here("figures/figure-s1.jpg"))
```

```{r figure-s2, fig.cap="**Figure S2.** An example instruction sheet distributed to volunteer residents collecting point-of-use samples from sentinel homes."}
knitr::include_graphics(here::here("figures/figure-s2.png"))
```

```{r figure-s3, fig.cap="**Figure S3. (a)** In the full (Zones 1 and 2) model, local multi-year smooths capturing the deviations of each series from the global and location-specific smooths. **(b)** Coagulation pH at the treatment plant supplying Zone 2."}
knitr::include_graphics(here::here("figures/figure-s3.png"))
```

```{r figure-s4, fig.cap="**Figure S4.** In the Zone 1 model, local multi-year smooths capturing the deviations of each series from the global and seasonal smooths."}
knitr::include_graphics(here::here("figures/figure-s4.png"))
```

```{r figure-s5, fig.cap="**Figure S5.** Total aluminum in pipe rack effluent."}
knitr::include_graphics(here::here("figures/figure-s5.png"))
```