### 2023-05-16 README.Rmd
update readme with preprint DOI

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

# cct-monitoring

This repository contains the code and data necessary to reproduce the main results of the preprint ["Sentinel pipe racks quantify orthophosphate's effect on lead release into drinking water"](https://doi.org/10.26434/chemrxiv-2023-h1459).