### 2023-08-15 Rmarkdown/paper-si.Rmd
revisions post-review

---
title: "*Supplementary information for:* Evaluating sentinel pipe racks for monitoring lead release and optimizing corrosion control"
author: "Benjamin F. Trueman^§,^\\*, Javier A. Locsin^§^, Wendy H. Krkošek^†^, and Graham A. Gagnon^§^"
date: "^§^Centre for Water Resources Studies, Department of Civil & Resource Engineering, Dalhousie University, 1360 Barrington St., Halifax, Nova Scotia, Canada B3H 4R2  \n&nbsp;  \n^†^Halifax Water, 450 Cowie Hill Rd., Halifax, Nova Scotia, Canada,  B3P 2V3  \n&nbsp;   \n&nbsp;*Corresponding author  \n&nbsp;E-mail: benjamin.trueman@dal.ca    \n&nbsp;Tel:902.494.6070  \n&nbsp;Fax: 902.494.3105"
output:
  word_document:
    reference_docx: StylesTemplate.docx
bibliography: references.bib
csl: environmental-science-and-technology.csl
---

```{r setup, include=FALSE}
here::i_am("Rmarkdown/paper.Rmd")
knitr::opts_chunk$set(echo = FALSE, dev = "png", dpi = 600)
options(dplyr.summarise.inform = FALSE)
source(here::here("R/03-inform-prior-ml.R"))
source(here::here("R/04-model-ml.R"))
library("posterior")
library("here")
wq <- read_csv(here("data-clean/summary-table.csv"))
wq_si <- read_csv(here("data-clean/summary-table-si.csv"))
```

```{r functions}
# this is the opposite of fill()

unfill_vec <- function(x, placeholder = "") {
  same <- x == dplyr::lag(x)
  ifelse(!is.na(same) & same, placeholder, as.character(x))
}

```

This document has X pages, X figures, and X tables.

```{r summary-wq-si}
options(knitr.kable.NA = "")
wq_si %>% 
  group_by(param) %>% 
  mutate(unit = unfill_vec(unit)) %>% 
  ungroup() %>% 
  transmute(
    Parameter = unfill_vec(param),
    Unit = str_replace(unit, "ug", "µg"),
    Zone = zone,
    Median = round(med, 1), 
    `Lower quartile` = round(lq, 1),
    `Upper quartile` = round(uq, 1)
  ) %>% 
  knitr::kable(caption = "**Table S1.** Summary of water quality in pipe rack effluent, by zone.")

```

```{r figure-s1, fig.cap="**Figure S1.** An example of the pipe racks installed in Zones 1 and 2."}
knitr::include_graphics(here::here("figures/figure-s1.jpg"))
```

```{r figure-s2, fig.width=6.5, fig.height=5, fig.cap="**Figure S2.** X-ray diffraction patterns characterizing lead pipe corrosion scale recovered from sites 1--3 (a--c, respectively), by layer (L1--L3/4)."}
knitr::include_graphics(here::here("figures/figure-s2.png"))
```

```{r figure-s3, fig.cap="**Figure S3.** An example instruction sheet distributed to volunteer residents collecting point-of-use samples from sentinel homes."}
knitr::include_graphics(here::here("figures/figure-s3.png"))
```

```{r figure-s4, fig.cap="**Figure S4. (a)** In the full (Zones 1 and 2) model, local multi-year smooths capturing the deviations of each series from the global and location-specific smooths. **(b)** Coagulation pH at the treatment plant supplying Zone 2."}
knitr::include_graphics(here::here("figures/figure-s4.png"))
```

```{r figure-s5, fig.cap="**Figure S5.** In the Zone 1 model, local multi-year smooths capturing the deviations of each series from the global and seasonal smooths."}
knitr::include_graphics(here::here("figures/figure-s5.png"))
```

```{r figure-s6, fig.cap="**Figure S6.** Total aluminum in pipe rack effluent."}
knitr::include_graphics(here::here("figures/figure-s6.png"))
```

# Lead pipe scale analysis

This section describes our procedure for identifying crystalline phases in three lead pipe samples recovered from three sites in Zone 2.

## Scale extraction

Lead pipe samples were excavated carefully, sealed at either end, and delivered to the laboratory. We cut a 0.2 m long section from each sample with a hacksaw, and we cut each section in half longitudinally to reveal the interior. Pipe interiors were photographed, and corrosion scale was collected with a stainless steel spatula in discrete layers, differentiated by colour, texture, and distance from the pipe wall. Pipe samples yielded distinct outer (L1), middle (L2), and inner (L3--L4) layers. The collected scale was finely ground with a mortar and pestle, which was cleaned with dilute nitric acid and pure water between samples.

## X-ray diffraction

We identified crystalline phases in the scale samples using a Rigaku Ultima IV X-ray diffractometer with a copper K$\alpha$ radiation source. The X-ray tube was operated at 35 kV and 30 mA. Scans were acquired over the range 10 - 70&deg; (2$\theta$) with a step size of 0.04&deg; and a scan speed of 0.8 &deg; min^-1^. Diffraction patterns were compared against the standard powder diffraction files listed in Table S2.

```{r table-s2}

tibble::tribble(
  ~Phase, ~`PDF entry no.`,
  "Chloropyromorphite", "96-900-4202",
  "Hydroxypyromorphite", "00-086-0236",
  "Hydrocerussite", "96-201-2741",
  "Cerussite", "00-047-1734",
  "Litharge", "00-085-1739",
  "Lead", "96-900-8478"
) %>% 
  knitr::kable(caption = "**Table S2.** Standard X-ray diffraction patterns and their powder diffraction file entry numbers.")

```

## Results

Corrosion scale in the first pipe was highly crystalline (Figure 3a). The top two scale layers were dominated by a pyromorphite mineral, and the bottom layer contained hydrocerussite and pyromorphite. While the best-matching pyromorphite mineral was chloropyromorphite (Pb~3~(CO~3~)~2~(OH)~2~), standard diffraction patterns for chloro- and hydroxyl- pyromorphites are quite similar, and our match to chloro-pyromorphite is tentative. We identified it based on two key differences: 1) between 29.5 and 30.5 degrees (2$\theta$), chloropyromorphite yields two peaks to hydroxypyromorphite's one, and between 45 and 47 degrees, chloropyromorphite yields three clearly separated peaks to hydroxypyromorphite's two [@lytle_inhibition_2009].

Pipes extracted at the other two sites 2--3 had similar corrosion scale to that of site 1 (Figure 3b--c, respectively). A pyromorphite mineral was present, although the match to chloropyromorphite was evenless certain here than in the XRD patterns representing site 1. XRD patterns also matched the lead carbonates hydrocerussite (sites 2--3) and cerussite (site 2), as well as the lead oxide litharge (site 3).

# References